---
title: "Student Performance ML"
output: html_notebook
---

https://archive.ics.uci.edu/ml/datasets/Student+Performance
1 school - student's school (binary: 'GP' - Gabriel Pereira or 'MS' - Mousinho da Silveira)
2 sex - student's sex (binary: 'F' - female or 'M' - male)
3 age - student's age (numeric: from 15 to 22)
4 address - student's home address type (binary: 'U' - urban or 'R' - rural)
5 famsize - family size (binary: 'LE3' - less or equal to 3 or 'GT3' - greater than 3)
6 Pstatus - parent's cohabitation status (binary: 'T' - living together or 'A' - apart)
7 Medu - mother's education (numeric: 0 - none, 1 - primary education (4th grade), 2 - 5th to 9th grade, 3 - secondary education or 4 - higher education)
8 Fedu - father's education (numeric: 0 - none, 1 - primary education (4th grade), 2 - 5th to 9th grade, 3 - secondary education or 4 - higher education)
9 Mjob - mother's job (nominal: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')
10 Fjob - father's job (nominal: 'teacher', 'health' care related, civil 'services' (e.g. administrative or police), 'at_home' or 'other')
11 reason - reason to choose this school (nominal: close to 'home', school 'reputation', 'course' preference or 'other')
12 guardian - student's guardian (nominal: 'mother', 'father' or 'other')
13 traveltime - home to school travel time (numeric: 1 - <15 min., 2 - 15 to 30 min., 3 - 30 min. to 1 hour, or 4 - >1 hour)
14 studytime - weekly study time (numeric: 1 - <2 hours, 2 - 2 to 5 hours, 3 - 5 to 10 hours, or 4 - >10 hours)
15 failures - number of past class failures (numeric: n if 1<=n<3, else 4)
16 schoolsup - extra educational support (binary: yes or no)
17 famsup - family educational support (binary: yes or no)
18 paid - extra paid classes within the course subject (Math or Portuguese) (binary: yes or no)
19 activities - extra-curricular activities (binary: yes or no)
20 nursery - attended nursery school (binary: yes or no)
21 higher - wants to take higher education (binary: yes or no)
22 internet - Internet access at home (binary: yes or no)
23 romantic - with a romantic relationship (binary: yes or no)
24 famrel - quality of family relationships (numeric: from 1 - very bad to 5 - excellent)
25 freetime - free time after school (numeric: from 1 - very low to 5 - very high)
26 goout - going out with friends (numeric: from 1 - very low to 5 - very high)
27 Dalc - workday alcohol consumption (numeric: from 1 - very low to 5 - very high)
28 Walc - weekend alcohol consumption (numeric: from 1 - very low to 5 - very high)
29 health - current health status (numeric: from 1 - very bad to 5 - very good)
30 absences - number of school absences (numeric: from 0 to 93)

31 G1 - first period grade (numeric: from 0 to 20)
31 G2 - second period grade (numeric: from 0 to 20)
32 G3 - final grade (numeric: from 0 to 20, output target)

"It is more difficult to predict G3 without G2 and G1, but such prediction is much more useful."

reading in data
```{r}
setwd("C:/Users/lujas/Desktop/Student Performance")
data = read.csv("X_train.csv")
y_train = read.csv("y_train.csv")
x_test = read.csv("X_test.csv")
head(data)
```

Factoring categorical variables (maybe one hot encoding later)
NOT numerical variables into factors (yet?)
```{r}
data$school = as.factor(data$school)
data$sex = as.factor(data$sex)
data$address = as.factor(data$address)
data$famsize = as.factor(data$famsize)
data$Pstatus = as.factor(data$Pstatus)
data$Mjob = as.factor(data$Mjob)
data$Fjob = as.factor(data$Fjob)
data$reason = as.factor(data$reason)
data$guardian = as.factor(data$guardian)
data$schoolsup = as.factor(data$schoolsup)
data$famsup = as.factor(data$famsup)
data$paid = as.factor(data$paid)
data$activities = as.factor(data$activities)
data$nursery = as.factor(data$nursery)
data$higher = as.factor(data$higher)
data$internet = as.factor(data$internet)
data$romantic = as.factor(data$romantic)

```

Check unique
Check missing
```{r}
table(is.na(data)) # no missing values
table(duplicated(data)) # no duplicates
```

Look at distribution of quantitative variables
```{r}
library(reshape2)
library(ggplot2)
melted_data <- melt(data[,-1])      #no StudentID variable for graphic
ggplot(melted_data, aes(x=value)) +
  geom_histogram(bins=10) +
  facet_wrap(~variable, scales="free_x")

mylist <- mapply(table, data[,-1])
mylist
```


Creation of new variables to better measure the dataset

Good visualizations
Trying to predict good final score (g3)

Need to one-hot encode variables
Use x-train, y-train (has g3)
then x-test to compare y-test

logistic, RF, 
```{r}

```

XGBoost
```{r}
install.packages("xgboost")
library(xgboost)
library(Matrix)

dim(data)
dim(y_train)

sparse_matrix = sparse.model.matrix( ~ .-1, data = data) #one hot encoding of categoricals
output_vector = y_train[,-1]

bst <- xgboost(data = sparse_matrix, label = output_vector, max_depth = 4,
               eta = 1, nthread = 2, nrounds = 10,objective = "count:poisson")
```

```{r}
importance <- xgb.importance(feature_names = colnames(sparse_matrix), model = bst)
head(importance)
importanceRaw <- xgb.importance(feature_names = colnames(sparse_matrix), model = bst, data = sparse_matrix, label = output_vector)
# Cleaning for better display
importanceClean <- importanceRaw[,`:=`(Cover=NULL, Frequency=NULL)]
head(importanceClean)
xgb.plot.importance(importance_matrix = importance)
```

Comparing
Should create metric of absolute value of difference to test model efficiency
```{r}
test_matrix = sparse.model.matrix( ~ .-1, data = x_test) #one hot encoding of categoricals
predictions = predict(bst, newdata = test_matrix)

y_test = read.csv("y_test.csv")
aaa = cbind(y_test, predictions)
View(aaa)

aaa$diff = aaa$predictions - aaa$G3

hist(aaa$diff, breaks = seq(-8,11,by = 1))
summary(aaa$diff)
```

